>

### 2025-01-29 write-ups/kid_tangrams_cogsci_2025/kid_tangrams_cogsci_2025.Rmd
final pass of edits from Yannick and mike, and length futzing

---
title: "Preschoolers form conventional pacts with each other to communicate about novel shapes"
bibliography: preschool-tangrams.bib
csl: apa7.csl
document-params: "10pt, letterpaper"

@@

abstract: >
   Learning language requires learning not only the content of language, but also skills that allow us to use language to communicate. Iterated reference games provide a window into such skills, requiring rich communication as participants jointly converge on mutually understandable names for initially novel objects. Some classical experiments with young children are interpreted as showing that 4-5-year old children are incapable of succeeding at iterated reference games. Here, we revisit the question of young children's referential communicative abilities using a simpler, child-friendly paradigm. Across 51 pairs of 4-5-year old children, we find that preschool-aged children were successful in establishing reference. Children were 85% accurate, and they often used descriptions similar to their partner's. These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.

@@

Learning a language requires learning not only the content of that language, but also how to use the language to communicate. One test environment for language use is referential communication, the ability to describe a target so an interlocutor can pick it out from a set of possibilities. Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the description they provide to their beliefs about the interlocutor's knowledge state.

One test of referential communication skill is *iterated reference games* where one describes abstract shapes from an array of images repeatedly to the same partner [@clark1986; @krauss1964; @hawkins2020b; @boyce2024]. Over repetition, features of the initial verbose descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image. Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform referring expressions in later rounds. 
Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.

A classic study suggests that 4-5-year old preschoolers struggle with child-child referential communication [@glucksberg1966]. In their paradigm, one child was given a set of 6 blocks in a specific order. Their task was to describe the images on the block so their partner could pick out the corresponding block. As children described and selected blocks, they stacked them on pegs. While 4-5-year old children were successful on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had drawings of abstract shapes and there was no visual access. Even after multiple rounds with the same images, children were not able to correctly order the blocks. 
Similar experiments with older children indicated a gradual improvement through adolescence both for initial accuracy and for the increase in accuracy across repetitions. Still, even the 9th grade sample was noticeably worse than the sample of adult college students [@glucksberg1967; @krauss1969]. That the task was difficult even among teenagers suggests that the complexity of the stacking paradigm and the large number of potential targets present significant task demands that might prevent accurate measurement of children's abilities.

```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure. Panel A shows the experimental setup with the teller and guesser across the table from each other. Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only. Panel C shows the procedure for Experiment 1; within critical blocks targets were ordered randomly. Panel D shows the procedure for Experiment 2. \\label{game}", cache=FALSE}

@@

More recently, a number of studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6 year olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009], showing that children at this age can reason about others’ perspectives to more efficiently communication.

While children are sensitive to others’ perspectives, they still seem to struggle with appropriately tailoring the specificity of their utterances to the visual context, often resulting in utterances that are under-informative [@matthews2012; @leung2024].  When children must describe one of two very similar images, 4 and 5-year olds struggle to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019]. By 5-years old, children are sensitive to utterances that are over- or under-informative, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].

A common thread between many studies in referential expression production is that children perform better when the cognitive load of the task is reduced, i.e. when there are fewer possible referents to consider [@abbot-smith2016]. This paired with evidence of emerging communicative skills in young children suggests that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.

Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game. Recent work shows that 8-10-year olds exhibit adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance  and highly variable between dyads [@branigan2016]. 4-6-year olds have also been shown to rely on conventionalized gestures with their partner in an iterated reference game where children could only use gestures to communicate [@bohn2019]. Recent work has established that young children can succeed at iterated reference games when playing with their parents [@leung2024]. 4-8 year olds' played an iterated reference game with a parent using a simple, child-friendly tablet-based task. In this task, even 4-year olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024]. Together, these findings provide evidence of young children’s ability to successfully communicate about novel referents under the right conditions.

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across 51 pairs of 4-5-year old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

4 and 5-year old children were recruited from a university preschool during the school day. Children played with another child from the same class. Experiment 1 was conducted between June and August 2023. Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials. We had 19 complete games and 1 included incomplete game. Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.

### Materials
For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B). We coded the matching game using Empirica and hosted it on a lab server [@almaatouq2020]. We then accessed the game on the web on tablets that were locked in a kiosk mode so children could not navigate away from the game.

### Procedure
Once a pair of children agreed to play the game, a research assistant took them to a quiet testing room. Children were introduced to a stuffed animal "Smurfy" who wanted to play a matching game. Children sat across a table from each other, each with a tablet in front of them (Figure \ref{game}A). On each trial, one child was the "teller" and saw a black box around one of two images on their screen and was asked to "tell Smurfy what they see" in the black box. The "guesser" saw the same two images in a randomized order and tried to select the described image to help Smurfy make a match. When the guesser selected an image, both children received feedback in form of a smiley or frowny face and an audible sound. After each trial, children switched roles. Children passed Smurfy back and forth to help them keep track of whether they were the "guesser" or "teller" on a given trial.

@@

Children's selections and the time to selection were recorded from the experiment software. Children's descriptions were transcribed from the video using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters. Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used. 
We excluded trials where the "teller" did not produce a description, or where all description was unintelligible and impossible to transcribe. After exclusions, we had `r clean_2 |> filter(!is.na(repNum)) |> nrow()` trials remaining.

@@

## Results

### Accuracy and speed

@@

Our primary measure of interest was whether children could accurately communicate the intended target. <!-- as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966]. -->  As shown in Figure \ref{acc}, children were generally accurate in their selections. To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`). This level of accuracy is generally in-line with accuracies from 4-year-olds playing with their parents in @leung2024 and indicates that children can understand and succeed at the task.

@@

As another measure of children's performance, we looked at how long each trial took to see if children were getting faster over time. We ran a Bayesian mixed effects linear regression predicting the time to selection in seconds.^[`r speed_form |> form()`] The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)` seconds / trial). This is a measure of efficiency as children were able to achieve the same accuracy levels in less time.

@@

We compared the semantic similarities between descriptions of the same target based on who produced the description (Figure \ref{sim}).  We used a Bayesian mixed effects linear regression to predict similarity.^[`r sim_across_between_form |> form()`] Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`). Utterances were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_1_results |> stats_text(3,3)`), which is expected since children are likely to be fairly consistent with themselves. However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions.

@@

In the first experiment, we adapted the paradigm of @leung2024 for pairs of children, taking an already simple set up and making it shorter. Our goal was to see if young children were at all able to provide adequate descriptions, so children received a lot of scaffolding around the experimental interaction. Sometimes, this included experimenters echoing children's descriptions which could potentially influence children's responses. In Experiment 2, we repeated the same paradigm, with a tighter experimental script and a larger sample size.

@@

### Participants
Experiment 2 was run between March and August of 2024, at the same university preschool as Experiment 1. No children participated in both experiments. 30 pairs of children completed all 16 critical trials, and 1 pair of children completed between 8 and 16 critical trials. Our target age range was 4 and 5 year olds, but one older 3-year-old was unintentionally included. Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.

@@

### Procedure
The procedure was much the same as Experiment 1(Figure \ref{game}D). We added an initial "bubble popping" exercise to give children practice with how to tap the tablet appropriately (this was an issue for some children in Experiment 1). The experimental script was fully written out and memorized by experimenters so children all received the same instructions. We wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections. Experimenters helped with game mechanics such as whose turn it was to tell and who should press the screen to move the game along, but avoided contributing or repeating any content about the images or the descriptions.

@@

In Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and accuracy was relatively stable over the course of the game (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}). 
The first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)` seconds / trial). Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens. Taken together, we find more evidence that children can successfully communicate with each other about these abstract shapes, and do so with increasing efficiency.

@@

The average length of descriptions on the first trial was `r stats_text(description_mod_2, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_2, 2)` words / trial, Figure \ref{length}). This is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.

@@

As a course measure of partner-sensitivity, we repeated the semantic analysis from Experiment 1. Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_2_results |> stats_text(3,3)`).

As Experiment 2 had 4 blocks, we examined whether descriptions were converging semantically toward the final description. We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`). Descriptions were more similar to each other if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`). Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant: `r sim_to_next_2_results |> stats_text(2,3)`.

As each partnership converged to their shared nicknames, partnerships often diverged from one another as groups focused on distinct aspects of the image. We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`), which indicates that games are converging to different conventions. These patterns of increasing similarity within games and increasing divergence between games match the patterns found for adults, albeit weakly [@boyce2024].

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>
  ggplot(aes(as.character(earlier), sim, color = target1)) +
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
  facet_wrap(~same_speaker) +
  scale_color_solarized() +
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
  theme(legend.position = "bottom",
            strip.background = element_blank(),
        panel.grid=element_blank()
)
```

@@

As the two experiments were similar to one another, we re-ran models pooling the data across the two experiments, using experiment number as a random effect. Pooling the two experiments, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_meta,1)`) and accuracy slightly, but unreliably, increased over the course of the game (OR of one trial later: `r stats_text(acc_mod_meta, 2)`). 
Descriptions produced by tellers averaged `r stats_text(description_mod_meta, 1)` words on the first trial, and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
Utterances were more similar if they came from the same partnership (increase in cosine similarity: `r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).

@@

We might expect that whether a description was successful influences whether the same description, or a variant of it, is employed in future rounds. Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt. To test this idea that accuracy might be predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar with the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller. This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful. <!--Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`. -->

@@

Prominent classical findings have been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, an emerging pattern of research finds that  young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. -->We re-addressed the question of preschoolers's performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

@@

Unlike adults, Children did not display an increase in accuracy nor a reduction in the length of referential expressions over the course of the game. This might be taken to mean that children are not adapting their descriptions over time. Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that childen are adapting their descriptions as the game unfolds. These null findings are likely a result of initially high accuracy in the first block and initially short utterances that have less possibility for reduction when compared to adult utterances.

It is unclear to what extent the uniformly short descriptions are a product of the simplified task or children's behavioral differences from adults. In this case, the low number of options and relatively easy to describe shapes may have obviated the need for long initial descriptions. Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024]. However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification. Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.

>>>>

---
title: >
  Preschoolers can form conventional pacts with each other 
  to communicate about novel referents
bibliography: preschool-tangrams.bib
csl: apa7.csl
document-params: "10pt, letterpaper"

@@

abstract: > 
  Learning language requires learning not only the content of language, but also how to use language to communicate. Iterated reference games provide a window into such skills, requiring rich communication as participants converge on mutually understandable names for initially novel referents. Some early experiments are interpreted as evidence that 4-5-year old children cannot converge to the mutually understandable names needed to communicate in an iterated reference game. Here, we revisit young children's referential communicative abilities using a simpler, child-friendly paradigm. Across 51 pairs of children, we found that 4-5 year olds successfully established reference with each other. Children were 85% accurate, and they often used descriptions similar to their partner's. These findings suggest that children’s capacity to construct effective referring expressions in novel contexts emerges earlier than once thought, consistent with the view that children show early pragmatic competence in supportive contexts.

@@

Learning a language requires learning not only the content of that language, but also how to use the language to communicate. One case study for language use is referential communication, the ability to describe a target so an interlocutor can pick it out from a set of possibilities. Adults show sensitivity to both the visual context and their audience during referential communication, calibrating the description they provide to their beliefs about the interlocutor's knowledge state.

Iterated reference games provide an important paradigm for studying referential communication. In these games, one player repeatedly describes a set of abstract shapes to a partner so they can identify the target images [@clark1986; @krauss1964; @hawkins2020b; @boyce2024]. Over repetition, features of the initial descriptions are conventionalized as each pair comes to agree on a shared understanding of how to label each image. Success at this task requires mastery of a number of linguistic and communicative skills, including producing adequate initial descriptions, monitoring for comprehension, asking for clarification, and appropriately using the shared conversation history to inform later referring expressions. 
Studying how children play iterated reference games can provide insight into the developmental trajectory of the ability to produce referential expressions in order to achieve joint understanding.

One influential early study suggested that 4-5-year old preschoolers struggle with child-child referential communication [@glucksberg1966]. In their paradigm, one child was given a set of 6 blocks in a specific order. Their task was to describe the images on the block so their partner could pick out the corresponding block. As children described and selected blocks, they stacked them on pegs. While 4-5-year old children succeeded on practice trials with familiar shapes and visual access to each other's blocks, children failed on critical trials where the blocks had abstract drawings and there was no visual access. Even after multiple rounds with the same images, children were not able to correctly order the blocks. @glucksberg1966 attributed children's communicative failures to their production of ego-centric descriptions that did not account for the other child's perspective.

Similar experiments with older children indicated a gradual improvement through adolescence both for initial accuracy and for the increase in accuracy across repetitions. Still, even the 9th grade sample was noticeably worse than the adult college student sample [@glucksberg1967; @krauss1969]. Given that even teenagers had difficulties with the task, the complex stacking paradigm and the large number of potential targets may have posed task demands preventing accurate measurement of children's abilities.

```{r interface, fig.env = "figure*", fig.pos = "t!", out.width="\\textwidth", fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "Experimental setup and procedure. Panel A shows the experimental setup. Panel B shows the 4 possible targets; names for targets are for cross-reference with later figures only. Panel C shows the procedure for Experiment 1; within critical blocks targets were ordered randomly. Panel D shows the procedure for Experiment 2. \\label{game}", cache=FALSE}

@@

More recently, though no evidence has directly contradicted this early study, a number of other studies have revealed early emerging skills in the preschool years that support the use of referential communication. Preschool-aged children are faster to select a target when it is referred to in a consistent way [@graham2014; @matthews2010] and 6 year olds are more likely to use consistent referential expressions when their partner is consistent [@koymen2014], suggesting that young children are sensitive to the norm of consistent descriptions in cooperative communication. Further, preschool-aged children adapt the informativeness of their referential expressions based on the visual content available to their interlocutor [@matthews2006;@nadig2002;@nilsen2009]. By age 5, children integrate information about other's perspectives into their comprehension of utterances [@sanjuan2015]. Overall, by late preschool, children can can reason about others’ perspectives to communicate more effectively.

While children are sensitive to others’ perspectives, they still struggle to appropriately tailor the specificity of their utterances to the visual context, often resulting under-informative utterances [@matthews2012; @leung2024].  When children must describe one of two very similar images, 4 and 5-year olds sometimes neglect to mention the relevant features, although they do better when playing in an interactive game than when not [@grigoroglou2019]. By 5-years old, children are sensitive to over- and under-informative utterances, asking for clarification on some under-informative utterances and taking longer to make selections on over-informative utterances [@morisseau2013].

A common thread among many developmental studies is that children perform better when the cognitive load of the task is reduced, i.e. when there are fewer possible referents [@abbot-smith2016; @sanjuan2015].  These findings paired with evidence of emerging communicative skills in young children suggests that tailored referential expression production is possible in children, but it is likely to be masked when the task demands are too high.

Since @glucksberg1966, few studies have revisited the question of whether children can successfully communicate in an iterated reference game. In one study, 8-10-year olds exhibited adult-like patterns of increasing accuracy, increasing speed, and shorter descriptions across repetitions, but children’s accuracy was still far below adult performance  and highly variable between dyads [@branigan2016]. 4-6-year olds successfully used conventionalized gestures with their partners in an iterated reference game where children could only use gestures to communicate [@bohn2019]. 4-8 year old children succeeded at playing an iterated reference game with a parent using a simple, child-friendly tablet-based task [@leung2024]. In this task, even 4-year olds had an initial accuracy above 80%, which rose to above 90% in later repetitions [@leung2024]. Together, these findings provide evidence of young children’s ability to communicate about novel referents under the right conditions.

Given the task demands in @glucksberg1966 and work showing that children can succeed in a less demanding paradigm, here we revisit the question of child-child referential communication. In the present study, we re-examine young children's ability to establish effective referring expressions with each other in an iterated reference game using a simplified tablet-based paradigm. Across an initial study and a recplication including a total of 51 pairs of 4-5-year old children, we found that preschool-aged children were successful in an iterated reference game, suggesting that children's capacity to construct effective referring expressions in novel contexts emerges earlier than once claimed.

@@

4 and 5-year old children were recruited from a university preschool during the school day. Children played with another child from the same class. Experiment 1 was conducted between June and August 2023. Pairs of children were included in analyses if they completed at least 8 of the 12 critical trials. We had 19 games that completed 12 critical trials, and 1 game that completed 11 critical trials. Of the 40 children, 21 were girls, and the median age was 57 months, with a range of 48-70 months.

### Materials
For the target stimuli, we used four of the ten tangram images from @leung2024, chosen based on visual dissimilarity (Figure \ref{game}B). We coded the matching game using Empirica [@almaatouq2020], hosted it on a server, and then accessed the game on ablets that were locked in a kiosk mode so children could not navigate away from the game.

### Procedure
Once a pair of children agreed to play the game, a research assistant took them to a quiet testing room. Children were introduced to a stuffed animal "Smurfy" who wanted to play a matching game. Children sat across a table from each other, each with a tablet in front of them (Figure \ref{game}A). On each trial, one child was the "teller" and saw a black box around one of two images on their screen and was asked to "tell Smurfy what they see" in the black box. The "guesser" saw the same two images in a randomized order and tried to select the described image to help Smurfy make a match. When the guesser selected an image, both children received feedback in form of a smiley or frowny face and an excited or disappointed sound. After each trial, children switched roles. Children passed Smurfy back and forth to keep track of whether they were the "guesser" or "teller" on a given trial.

@@

Children's selections and the time to selection were recorded from the experiment software. Children's descriptions were automatically transcribed from the video using Whisper [@radford2022] for the first pass and then hand-corrected by experimenters. Transcripts were hand-annotated for when each trial started, who said each line, and what referential descriptions were used. 
We excluded trials where the "teller" did not produce a description, or where all description was unintelligible and impossible to transcribe. After exclusions, we had `r clean_1 |> filter(!is.na(repNum)) |> nrow()` trials remaining.

@@

## Results
### Accuracy and speed

@@

Our primary measure of interest was whether children could accurately communicate the intended target. <!-- as prior work is often interpreted as indicating the children at kindergarten age cannot communicate about abstract shapes successfully [@glucksberg1966]. --> To test for changes in accuracy over time, we fit a Bayesian mixed effects logistic regression predicting accuracy.^[`r acc_form |> form()`] Children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_1,1)`), and their accuracy slightly increased over the game (OR of one trial later: `r stats_text(acc_mod_1, 2)`, Figure \ref{acc}). This level of accuracy is generally in-line with accuracies from 4-year-olds playing with their parents in @leung2024 and indicates that children can understand and succeed at the task.

@@

As another measure of children's performance, we looked at how long children spent on each trial. We ran a Bayesian mixed effects linear regression predicting the time to selection in seconds.^[`r speed_form |> form()`] The first critical trial averaged `r stats_text(speed_mod_1, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_1, 2)` seconds / trial). Children were able to achieve the same accuracy in less time, suggesting that they were becoming more efficient at completing the task.

@@

We compared the semantic similarities between descriptions of the same target based on who produced the description (Figure \ref{sim}).  We used a Bayesian mixed effects linear regression to predict similarity.^[`r sim_across_between_form |> form()`] Utterances were more similar if they came from the same partnership (`r sim_across_between_1_results |> stats_text(2,3)`). Utterances were slightly more similar still if they came from the same person (`r sim_across_between_1_results |> stats_text(3,3)`), which is expected since children are likely to be fairly consistent with themselves. However, children used descriptions that were much more similar to their partner's than to other children's (Figure \ref{sim}), indicating sensitivity to their partner's expressions.

@@

In Experiment 1, we adapted the paradigm of @leung2024 for pairs of children, taking an already simple set-up and making it shorter. Our goal was to see if young children were at all able to provide adequate descriptions, so children received a lot of scaffolding around the experimental interaction. Sometimes, this scaffolding included experimenters echoing children's descriptions, which could potentially influence children's responses. In Experiment 2, we repeated the same paradigm, with a tighter experimental script and a larger sample size.

@@

### Participants
Experiment 2 was run between March and August of 2024, at the same preschool as Experiment 1. No children participated in both experiments. 30 pairs of children completed all 16 critical trials, and 1 pair of children completed 10 critical trials. Our target age range was 4 and 5 year olds, but one older 3-year-old was unintentionally included. Of the 62 children, 30 were girls, and the children had a median age of 56 months, and a range of 45-69 months.

@@

### Procedure
The procedure was much the same as Experiment 1(Figure \ref{game}D). We added an initial "bubble popping" exercise to give children practice tapping the tablet appropriately (this was an issue for some children in Experiment 1). The experimental script was fully written out and memorized by experimenters so children all received the same instructions. We wrote up contingency statements that the experimenter could use to prompt children who were not giving descriptions or making selections. Experimenters helped with game mechanics such as whose turn it was to tell and who should press the screen, but avoided contributing or repeating any content about the images or the descriptions.

@@

In Experiment 2, children's accuracy was above chance (Odds Ratio: `r stats_text(acc_mod_2,1)`) and relatively stable over time (OR of one trial later: `r stats_text(acc_mod_2, 2)`, Figure \ref{acc}). 
The first critical trial averaged `r stats_text(speed_mod_2, 1)` seconds, and children got faster over time (`r stats_text(speed_mod_2, 2)` seconds / trial). Children were initially faster in Experiment 2 than Experiment 1, possibly due to the increased number of practice trials and pre-training on how to press the screens. Taken together, we find more evidence that children can successfully communicate with each other about these abstract shapes, and do so with increasing efficiency.

@@

The average length of descriptions on the first trial was `r stats_text(description_mod_2, 1)` words and description length was relatively stable over time (change of `r stats_text(description_mod_2, 2)` words / trial, Figure \ref{length}). This finding is comparable to Experiment 1, again finding that children produce short utterances without much change in length over time.

@@

As a coarse measure of partner-sensitivity, we repeated the semantic analysis from Experiment 1. Utterances were more similar if they came from the same partnership (`r sim_across_between_2_results |> stats_text(2,3)`) and were slightly more similar if they came from the same person (`r sim_across_between_2_results |> stats_text(3,3)`).

As Experiment 2 had 4 blocks, we examined whether descriptions were converging semantically toward the final description. We compared the utterances from the first three blocks to the descriptions in the last block using a Bayesian mixed effects linear regression predicting similarity.^[`r sim_to_last_form |> form()`] Over the first three blocks, descriptions became increasingly similar to the last block description (`r sim_to_last_2_results |> stats_text(2,3)`). Descriptions were more similar to each other if they came from the same child, which is expected as a sign of internal consistency (`r sim_to_last_2_results |> stats_text(3,3)`). Although over time descriptions did get more similar to the last block utterance, the semantic distance between adjacent block utterances was relatively constant (`r sim_to_next_2_results |> stats_text(2,3)`).

Partnerships often diverged from one another as groups focused on distinct aspects of the image. We tested whether descriptions in different games to the same target diverged over time using a Bayesian mixed effects linear regression.^[`r sim_across_form |> form()`] As the games progressed, descriptions from different games became slightly further apart in semantic space (`r sim_across_2_results |> stats_text(2,3)`), which indicates that games are converging to different conventions. These patterns of increasing similarity within games and increasing divergence between games qualitatively match the patterns found for adults [@boyce2024].

```{r sim2, fig.env="figure", fig.pos = "t", fig.align = "center", out.width="100%", fig.width=5, fig.height=3, fig.cap = "Semantic similarity between descriptions from earlier blocks (1-3) and the last block.  Heavy dots are means with bootstrapped 95% CIs; light dots are individual values. \\label{tolast}" }
sims_2 |>
  filter(game1 == game2) |>
  mutate(later = ifelse(block1 > block2, block1, block2), earlier = ifelse(block1 > block2, block2, block1)) |>
  filter(later == 4) |>
  mutate(same_speaker = ifelse(speaker1 == speaker2, "Same teller", "Different teller")) |>
  ggplot(aes(earlier, sim, color = target1)) +
    geom_smooth(color="black", method="lm")+
  geom_point(alpha = .1, position = position_dodge(width = .2)) +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2)) +
    #stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .2), geom="line", lty="dashed") +
  #  geom_smooth( method="lm")+
  facet_wrap(~same_speaker) +
  scale_color_solarized() +
  scale_x_continuous(breaks=c(1,2,3), labels=c("1", "2", "3"))+
  labs(x = "Comparing an earlier block (1-3) to last block (4)", y = "Cosine Similarity", color = "Target") +
  theme(legend.position = "bottom",
            strip.background = element_blank(),
        #panel.grid=element_blank()
)

@@

As the two experiments were similar to one another, we re-ran models pooling the data across the two experiments, using experiment number as a random effect. Pooling the two experiments, children's accuracy was above chance (OR: `r stats_text(acc_mod_meta,1)`) and accuracy numerically increased over the course of the game, although the credible interval included 0 (OR of one trial later: `r stats_text(acc_mod_meta, 2)`). 
Descriptions produced by tellers averaged `r stats_text(description_mod_meta, 1)` words on the first trial, and description length was relatively stable over time (change of `r stats_text(description_mod_meta, 2)` words per trial).
Utterances were more similar if they came from the same partnership (increase in cosine similarity: `r sim_across_between_mod_meta |> stats_text(2,3)`) and were slightly more similar still if they came from the same person within the partnership (`r sim_across_between_mod_meta |> stats_text(3,3)`).

@@

We might expect that whether a description was successful influences whether the same description, or a variant of it, is employed in future rounds. Intuitively, successful descriptions can be copied and built upon, while unsuccessful descriptions should be replaced by a fresh attempt. To test this idea that accuracy might be predictive of similarity to future descriptions, we ran a post-hoc Bayesian linear model predicting similarity to the next block description in terms of accuracy.^[`r form(sim_acc_to_next_form)`] Descriptions that elicited a correct response were more similar to the next block description (`r stats_text(sim_acc_to_next_mod_meta, 2, 3)`) with no substantial interaction with block number or whether both descriptions came from the same teller. This pattern of results is consistent with the expectation that children are more likely to stick to their own description or repeat the other child's description if it was previously successful. <!--Numerically, the size of the boost for being correct was similar to the size of the boost from coming from the same person `r stats_text(sim_acc_to_next_mod_meta, 6,3)`. -->

@@

Prominent early findings been interpreted as evidence that young children cannot overcome their ego-centrism to coordinate with each other in reference games [@glucksberg1966]. However, more recent developmental have found that young children show emerging communicative and pragmatic sensitivity, especially when cognitive demands are low. <!--Recent work has established that children are successful at producing and understanding referential expressions to novel shapes when playing reference games with their parents. -->We revisited the question of preschoolers's performance in child-child reference games, using a scaffolded paradigm to reduce extraneous task demands.

@@

Unlike adults, children did not display an increase in accuracy or a shortening of referential expressions over the course of the game. Still, our findings show that descriptions became increasingly similar to descriptions in the last block and that successful utterances were more similar to future utterances, suggesting that children are adapting their descriptions as the game unfolds. These null findings are likely a result of initially high accuracy in the first block and initially short utterances that leave little room for reduction.

It is unclear to what extent the uniformly short descriptions we observed are a product of the simplified task or children's behavioral differences from adults. In this case, the low number of options and relatively easy to describe shapes may have obviated the need for long initial descriptions. Indeed, adult controls in @leung2024 used shorter initial descriptions than adults in studies with larger arrays of harder to distinguish images [@hawkins2020b; @boyce2024]. However, young children may also struggle to produce longer descriptions, and young children may be more willing to take guesses when adults would seek additional clarification. Especially in light of other work suggesting that conceptual pact formation and reduction in utterance length sometimes decouple in adults [@boyce2024], further empirical work on the factors driving verbosity in reference games is warranted.

The generalizability of our results is limited by the target population, the target images, and the task structure. We sampled a convenience population of children at a university nursery school. The set of tangram images may be easier to distinguish and have higher codability than other target images used in adult reference games. We specifically targeted children's abilities to construct referring expressions that can be jointly understood, so children were provided scaffolding around taking turns and talking to their partner. Thus, children's performance should be taken as a proof-of-concept about ability, rather than a claim about how generally children spontaneously demonstrate these abilities.
